---
title: "SATools"
author: "Systems Analysis UK"
date: '2022-07-21'
output: html_document
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
## Load Libraries ####
packages <- list(
  "rhdf5",
  "DT",
  "magrittr",
  "fst",
  "tidyverse",
  "data.table",
  "Biostrings",
  "scales",
  "pbapply",
  "pbmcapply",
  "parallel",
  "rvest")
message("Loading Packages...", appendLF = FALSE)
suppressPackageStartupMessages(package.load.check <- sapply(packages, require, character.only = TRUE))
if (!all(package.load.check)) stop("Not all packages loaded")
message("DONE")

## Set some useful defaults
options(mc.cores=detectCores())    ## Allow the mc*apply functions to use all available cores
setDTthreads(threads = detectCores())
options(stringsAsFactors = FALSE)  ## Don't allow data.frames to use strings as factors by default
options(scipen = 999)
options(ignore.interactive = TRUE)

## Illumina Theme, use for ggplot2 graphics e.g.  ggplot(data, aes(x = x, y = y)) + theme_Illumina() + geom_point()
theme_Illumina <- function(){theme_grey(base_size = 24, base_family = "sans")}
theme_set(theme_Illumina())
### Functions #####
source("SAReportFunctions.R")
```

```{r loadInteropData, echo=FALSE, include=FALSE}
argv <- commandArgs(trailingOnly = TRUE)
in.dir <- argv[1]
runID <- in.dir %>% basename
## Just for development
#in.dir <- "/illumina/scratch/Lightning/Informatics/GoldenRuns/Galapagos/20220603_P2-13_0115_B12YCYOLO3/"
interop.file <- list.files(in.dir, pattern = "_fullDataWithIQS.csv", full.names = TRUE)
if (!is.null(interop.file)) {
interop <- fread(interop.file)
} else {
  stop("Can not find _fullDataWithIQS.csv file")
}

interop <- fread(interop.file)
```

```{r LoadCachedData, echo=FALSE, include=FALSE}
in.files <- list.files(file.path(in.dir, "Data", "InstrumentedData"),
                       recursive = TRUE, 
                       pattern = "AggregatedH5_s.*\\.csv", 
                       full.names = TRUE)

## Load in the cached CSV files ####
message("Found ", in.files %>%
length, " cache files to load")
results <- in.files %>% 
  pbmclapply(., fread) %>% 
  rbindlist(fill = TRUE)

## Post Processing ####
results[, onInThisChannel := factor(
onInThisChannel,
levels = c("Empty", "FALSE", "TRUE"),
labels = c("Empty", "Off", "On")
)]

results <- results[!is.na(cycle),]

cycles <- results$cycle %>% unique

## Variable Analysis ####
variable.of.interest <- "run_lane"
results[,var:=str_c(run,"_",lane) %>% str_remove("_ASampleFlowcell")]

results <- results %>%
  mutate(Read = ifelse(cycle <= 151, 'R2', 'R1'))
  
```

# `r runID`
### Summary
```{r generateLaneSummary, echo = FALSE}
summary <- results %>% 
  .[,.(
    `mapq` = `mapq` %>% mean(na.rm = TRUE) %>% round(2),
    `Qscore` = `Qscore` %>% mean(na.rm = TRUE) %>% round(2),
    `PF` = `PF` %>% mean(na.rm = TRUE) %>% round(2),
    `is.empty` = `is.empty` %>% mean(na.rm = TRUE) %>% round(2),
    `isize` = `isize` %>% median(na.rm = TRUE) %>% round(2)
    ),
    by = .(lane,tile)]
```

```{r outputSummaryHTMLtable, echo = FALSE}
datatable(summary, filter = 'top')
```

`r format(Sys.time(), '%d %B, %Y')`

## Analyses {.tabset}

### PRIMETA

#### Primary Metrics Analysis {.tabset}

Code used for the analysis: https://git.illumina.com/mgataric/quality_models

##### Quality Drivers All Cycles

Main drivers of quality deterioration over all cycles:

```{r Top_drivers, echo=FALSE, fig.cap="", out.width = '55%'}
plt = primeta.plot('var_bars.png')
if (is.character(plt)){
  knitr::include_graphics(plt)
}
``` 

```{r Drives_per_tile_group, echo=FALSE, fig.cap="", out.width = '100%'}
plt = primeta.plot('scores_and_ranks_per_Surface_Lane.png')
if (is.character(plt)){
  knitr::include_graphics(plt)
}
```

```{r Drivers_med_scores_and_coef, echo=FALSE, fig.cap="", out.width = '100%'}
plt = primeta.plot('multilev_reg_per_tile.png')
if (is.character(plt)){
  knitr::include_graphics(plt)
}
```

##### Quality Prediction

Predicted change in quality if a metric remained constant over all cycles:

###### - If the main quality driver is kept constant throughout run:
```{r Primeta_phasing, echo=FALSE, fig.cap="", out.width = '100%'}
plt = primeta.plot('pred_changed_key_driver.png')
if (is.character(plt)){
  knitr::include_graphics(plt)
}
```

###### - If Image Quality Score is kept constant throughout run:
```{r Primeta_FWHM_green, echo=FALSE, fig.cap="", out.width = '100%'}
plt = primeta.plot('pred_changed_Focus.png')
if (is.character(plt)){
  knitr::include_graphics(plt)
}
```

###### - If Pre/Phasing Weight is kept constant throughout run:
```{r Primeta_lock_phasing_weight, echo=FALSE, fig.cap="", out.width = '100%'}
plt = primeta.plot('pred_changed_PP.png')
if (is.character(plt)){
  knitr::include_graphics(plt)
}
```

##### Quality Drivers Last 10 Cycles

Main drivers of quality variation among different tiles in the last 10 cycles:

```{r varriation_last_cyc, echo=FALSE, fig.cap="", out.width = '55%'}
plt = primeta.plot('var_bars_last_cycles.png')
if (is.character(plt)){
  knitr::include_graphics(plt)
}
```

```{r scores_per_cyc, echo=FALSE, fig.cap="", out.width = '100%'}
plt = primeta.plot('scores_per_cycle_group.png')
if (is.character(plt)){
  knitr::include_graphics(plt)
}
```

```{r multi_reg_per_cyc, echo=FALSE, fig.cap="", out.width = '100%'}
plt = primeta.plot('multi_reg_per_cycle_group.png')
if (is.character(plt)){
  knitr::include_graphics(plt)
}
```

##### Metrics

Covariance of metrics used in analysis and their correlation to the response variable:

```{r Primeta_merged_corr_metrics, echo=FALSE, fig.cap="", out.width = '100%'}
plt = primeta.plot('corr_merged_metrics.png')
if (is.character(plt)){
  knitr::include_graphics(plt)
}
```

Breakdown of new metrics constructed from those that are highly correlated (new metrics are with + signs):

```{r Primeta_merged_metrics, echo=FALSE, fig.cap="", out.width = '100%'}
plt = primeta.plot('^merged_metrics.png')
if (is.character(plt)){
  knitr::include_graphics(plt)
}
```

#### End
```{r name, echo=FALSE, fig.cap="", out.width = '100%'}
#knitr::include_graphics(primeta.plot(''))

```





### SNR

#### Results {.tabset}

```{r, settings, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.height = 20, fig.width = 12)
```

##### By Number of Neighbours Off/Empty
Signal to noise over cycles and grouped by the number of neighbouring wells that are either off in the associated channel or empty.
```{r SNR_by_Neighbours,  echo=FALSE, fig.cap="A caption", fig.width=30, fig.height=15, out.width = '100%'}
plotSNRforVariable(variable.of.interest = "neighboursOffOrEmpty", y.space = 4, x.space = 10)
```

```{r SNR_by_Neighbours_per_lane,  echo=FALSE, fig.cap="A caption", fig.width=30, fig.height=15, out.width = '100%'}
plotSNRforVariable(variable.of.interest = "lane", y.space = 4, x.space = 10)
```

```{r, SNR_by_Neighbours_mean,  echo=FALSE, fig.cap="A caption", fig.width=5, fig.height=10, out.width = '100%'}
plotSNRforVariable(variable.of.interest = "neighboursOffOrEmpty", returnMean = TRUE, y.space = 4, x.space = 10)
```

##### By Surface
Signal to noise over cycles and grouped by surface.
```{r SNR_by_surface, echo=FALSE, fig.cap="A caption", out.width = '100%', fig.height=15}
### SNR By Surface ####
plotSNRforVariable(variable.of.interest = "surface")


```

##### By Prior Base
Signal to noise over cycles and grouped by the prior base call.
```{r SNR,  echo=FALSE, fig.cap="A caption", out.width = '100%', fig.width=30, fig.height=15}
### SNR by prior base ####
plotSNRforVariable(variable.of.interest = "priorBase")


```

### Insert Size


```{r chastity_by_isize,  out.width = '100%', fig.width=10, fig.height=5}
p <- results[channel == 'ch1' & (PF) & isize > 0 & mapq > 10] %>% select(isize, Qscore, Read) %>% mutate(cut_number = cut_number(isize, n = 10)) %>% group_by(cut_number) %>% mutate(mean = mean(Qscore), sd = sd(Qscore)) %>% select(cut_number, mean, sd, Read) %>% unique()

p2 <- results[
channel == "ch1" & (PF) & mapq > 10 & isize > 0,
.(mean = mean(chastity), sd = sd(chastity)),
by = .(cut_number(isize,10),Read)]

p2 %>% ggplot(aes(x = cut_number, y = mean, ymin = mean-sd, ymax = mean+sd)) +
  #      theme_Illumina() +
  #geom_pointrange(aes(col = Read)) +
  geom_point(aes(group = Read, col = Read), size = 3) +
  geom_line(aes(group = Read, col = Read), size = 2) +
  xlab("Insert Size") + ylab("Mean Chastity") +
  ggtitle(paste("mean Chastity by insert size")) +
  theme(axis.text = element_text(size = 10)) +
  theme(axis.text.x = element_text(size = 15, angle = 45, vjust = 0.9, hjust = 0.8))
```

```{r intensity_desity}
current.cycle=27
results[(PF) &  (cycle == current.cycle) & mapq > 10  & isize > 0,
corrected,
by = .(cut_number(isize, 10), Read, onInThisChannel, channel)] %>%
ggplot(aes(x = corrected)) + geom_density(aes(fill = Read), alpha = 0.5) +
facet_grid(cut_number~onInThisChannel+channel) +
ggtitle(paste(
"C",
current.cycle,
variable.of.interest,
"Corrected Intensity : Insert Size"
)) +
  theme(axis.text = element_text(size = 10)) +
  theme(axis.text.x = element_text(size = 15, angle = 45, vjust = 0.9, hjust = 0.8))
```


```{r mapq_by_isize, fig.width=10, fig.height=10}
x <-   results[
(PF) & mapq >10 & isize > 0 &
(cycle <= 150 | cycle >=302-151),
mean(Qscore),
by = .(cut_number(isize,10), cycle,var)]

x %>%
ggplot(aes(x = cycle, y = cut_number)) +
geom_tile(aes(fill = V1)) 
```

```{r error_rate_by_isize, fig.width=10, fig.height=5}
results[isize <= 1500 & channel == "ch1" & mapq > 10  & isize > 0 & (PF) &  (cycle == 27) &
(cycleMisMatch %in% c("-", "=", DNA_BASES)),
.(error.rate = (sum(cycleMisMatch != "=", na.rm = TRUE) / .N) *
100, N = .N, medianIsize = median(isize) %>% round(0)),
by = .(cut_number(isize, 10), Read)] %>%
  ggplot(aes(x = cut_number, y = error.rate)) +
  #      theme_Illumina() +
  geom_point(aes(col = Read)) +
  geom_line(aes(group = Read, col = Read)) +
  xlab("Insert Size") + ylab("Error Rate") +
  guides(col = guide_legend(title = variable.of.interest)) +
  ggtitle(paste(
  "C",
  current.cycle,
  variable.of.interest,
  "Human Error Rate : Insert Size")) +
    theme(axis.text = element_text(size = 10)) +
    theme(axis.text.x = element_text(size = 15, angle = 45, vjust = 0.9, hjust = 0.8))
```

```{r mean_chastity, fig.width=12, fig.height=12}

x.bins <- 50
y.bins <- 50
results[(PF) &
cycle == 27 &
channel == "ch1", .(mean = mean(chastity)), by = .(calledBase, cut_interval(X, x.bins, dig.lab=5), cut_interval(Y, y.bins, dig.lab=5))] %>%
ggplot(aes(x = cut_interval, y = cut_interval.1)) +
geom_tile(aes(fill = mean)) + xlab("X") + ylab("Y") +
facet_wrap( ~ calledBase) +
ggtitle(paste(variable.of.interest, " mean chasity c", current.cycle)) +
  theme(axis.text = element_text(size = 10)) +
  theme(axis.text.x = element_text(size = 3, angle = 45, vjust = 0.9, hjust = 0.8)) 
```


### Crosstalk
#### Coming Soon

```{r crosstalk_chastity_per_lane, fig.width=10, fig.height=5}

results[(PF) &
channel == "ch1" &
cycle == current.cycle ,
.(meanChastity = mean(chastity),
sdChastity = sd(chastity)),
by = .(neighboursOffOrEmpty, lane)] %>%
ggplot(aes(x = neighboursOffOrEmpty, y = meanChastity)) +
geom_line(aes(col = lane)) +
ggtitle(paste("C", current.cycle, variable.of.interest, "Chastity : CrossTalk")) 

```
```{r, fig.width=10, fig.height=5}
results[(PF) & channel == "ch1" & (cycle == 27) &
(cycleMisMatch %in% c("-", "=", DNA_BASES)),
.(error.rate = (sum(cycleMisMatch != "=", na.rm = TRUE) / .N) *
100, N = .N),
by = .(neighboursOffOrEmpty, lane)] %>% filter(error.rate > 0) %>%
ggplot(aes(x = neighboursOffOrEmpty, y = error.rate)) +
theme_Illumina() +
geom_point(aes(col = lane)) +
geom_line(aes(col = lane)) +
guides(col = guide_legend(title = variable.of.interest)) +
ggtitle(paste(
"C",
current.cycle,
variable.of.interest,
"Human Error Rate : CrossTalk"
))
```

### Signal Decay
#### Coming Soon

```{r chastity_over_cycles, fig.width=10, fig.height=5}

cycles.to.plot <- cycles[c(
(cycles - 25) %>% is_weakly_greater_than(0) %>% which() %>% extract2(1),
cycles %>% length %>% divide_by(3) %>% round(0),
cycles %>% length %>% divide_by(3) %>% round(0) %>% multiply_by(2),
cycles %>% length
)]

x <-results[
channel == "ch1" & (PF) &
cycle %in% cycles.to.plot,
.(mean=mean(chastity)),
by = .( run, lane, surface, swath, tile, idx,cycle)]
x[,mean2:= mean %>% round(2)]
x %>%
ggplot(aes(x = mean2)) +
geom_histogram(aes(y = stat(count) / sum(count),fill = lane), bins = 100) +
xlim(0.5,1) +
scale_y_continuous(labels = percent) +
guides(fill = guide_legend(title = variable.of.interest)) +
ylab("Percent of Clusters") + xlab("Mean Chastity per Cluster") +
facet_grid(cycle~lane, scales = "free_y")
```


### Background
#### Coming Soon
